---
title: "Homework 3"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(splines)
library(astsa)
library(magrittr)
```

#1

```{r}
viewerData <- read.table("Viewership.txt", header = TRUE)
viewerData$Viewers <- log(viewerData$Viewers)
ggplot(data = viewerData, mapping=aes(x=ShowNum, y=Viewers)) + geom_point()
pairs(viewerData)
```
#1
There appears to be a linear relationship between log(viewers) and ShowNum, but there appears to be a clear change in the basic scatter plot around ShowNum = 40, this could be cause to look at a time series.

#2

```{r}
viewer.lm <- lm(Viewers~ShowNum, data = viewerData)
viewer.ACF <- acf(resid(viewer.lm), lag.max=20)
ACF.dframe <- data.frame(Lag=viewer.ACF$lag, ACF=viewer.ACF$acf)
ggplot(data=ACF.dframe, aes(x=Lag, y=ACF)) + geom_col()
```

The acf of the residuals from the linear model shows that there is temporal correlation in the residuals and thus this should be accounted for in the model.
This temporal correlation means that past viewership affects viewership in the future, so people may be more likely to watch the next episode of a series for example.

#3

```{r}
key.sarima <- data.frame(c(0,0,0,0,1,0),
                         c(1,0,0,0,1,0),
                         c(2,0,0,0,1,0),
                         c(0,0,1,0,1,0),
                         c(0,0,2,0,1,0),
                         c(1,0,2,0,1,0),
                         c(2,0,1,0,1,0),
                         c(2,0,2,0,1,0),
                         c(1,0,1,0,1,0),
                         c(0,0,0,1,1,0),
                         c(1,0,0,1,1,0),
                         c(2,0,0,1,1,0),
                         c(0,0,1,1,1,0),
                         c(0,0,2,1,1,0),
                         c(1,0,2,1,1,0),
                         c(2,0,1,1,1,0),
                         c(2,0,2,1,1,0),
                         c(1,0,1,1,1,0),
                         c(0,0,0,0,1,1),
                         c(1,0,0,0,1,1),
                         c(2,0,0,0,1,1),
                         c(0,0,1,0,1,1),
                         c(0,0,2,0,1,1),
                         c(1,0,2,0,1,1),
                         c(2,0,1,0,1,1),
                         c(2,0,2,0,1,1),
                         c(1,0,1,0,1,1),
                         c(0,0,0,1,1,1),
                         c(1,0,0,1,1,1),
                         c(2,0,0,1,1,1),
                         c(0,0,1,1,1,1),
                         c(0,0,2,1,1,1),
                         c(1,0,2,1,1,1),
                         c(2,0,1,1,1,1),
                         c(2,0,2,1,1,1),
                         c(1,0,1,1,1,1))
X <- bs(x=viewerData$ShowNum, knots=45, degree=1, Boundary.knots= c(min(viewerData$ShowNum), max(viewerData$ShowNum)+10))
s <- 10
aics <- rep(x=NA, times = ncol(key.sarima))
bics <- rep(x=NA, times = ncol(key.sarima))
for(i in 1:ncol(key.sarima)) {
  my.ts.model <- sarima(viewerData$Viewers, p=key.sarima[1,i], d=key.sarima[2,i], q=key.sarima[3,i], P=key.sarima[4,i], 
                        D=key.sarima[5,i], Q=key.sarima[6,i], S=s, xreg=X, details=FALSE)
  aics[i] <- my.ts.model$AIC
  bics[i] <- my.ts.model$BIC
 
}                
key.sarima[,which(aics == min(aics))]
key.sarima[,which(bics == min(bics))]
```
Running a sarima model for each possible combination of values for p, q, P, and Q, keeping d = 0 and D = 1, the appropriate values are p = 2, q = 0, P = 0, and Q = 1. As this gives the best AIC and BIC values. SOt he best model is SARIMA(2,0,0,0,1,1) where s = 10.

#4


#5
```{r}
viewers.model <- sarima(viewerData$Viewers, p = 2, d = 0, q = 0, P = 0, D = 1, Q = 1, S = s, xreg = X, details = FALSE)

decorResid <- resid(viewers.model$fit)
decorResid.ACF <- acf(decorResid, lag.max=36)
decorResid.ACF.dframe <- data.frame(Lag=decorResid.ACF$lag, ACF=decorResid.ACF$acf)

ggplot(data=decorResid.ACF.dframe, aes(x=Lag, y=ACF)) + geom_col()

fitSarima <- viewerData$Viewers - decorResid
ggplot(mapping=aes(x=fitSarima, y=decorResid)) + geom_point()

hist(decorResid)

```
Linearity is validated because the fitted values against the residuals plot appears to be linear, there appears to be independence as the ACF appears to no longer be temporally correlated, normality is seen through the normality of the residuals. And there is assumed to be equal variance based on our model having equal variance.

#6
```{r}

newX <- bs(viewerData$ShowNum, knots = 45, degree = 1)
train.set <- viewerData[1:(nrow(viewerData)-10),]
test.set <- viewerData[-(1:(nrow(viewerData)-10)),]
predicts <- max(viewerData$ShowNum) + seq(1, 10, by=1)
x.train <- X[1:(nrow(viewerData)-10),]
x.test <- X[-(1:(nrow(viewerData)-10)),]


viewer.for <- sarima.for(train.set$Viewers, p=2, d=0, q=0, P=0, D=1, Q=1, S=10, xreg = x.train, n.ahead=10, newxreg=x.test)
n <- 5


(viewer.for$pred - test.set$Viewers)^2 %>% mean() %>% sqrt()


low <- viewer.for$pred - qt(0.975, df=nrow(train.set)-n)*viewer.for$se
upr <- viewer.for$pred + qt(0.975, df=nrow(train.set)-n)*viewer.for$se
viewer.for$pred
low
upr
viewerData$Viewers[61:70]
```

This model results in an RPMSE of 0.1835391, with increases in viewership of a couple of percent between each show. The predictions for each show have a 95% confident lower estimate of  "2.114273 2.017491 2.017702 2.011621 2.020206 1.939456 1.908454 1.967351 1.926863 2.024952" and an upper estimate of "2.429220 2.391132 2.442838 2.472545 2.508832 2.449351 2.434875 2.506565 2.475839 2.581355", with the predictions being "2.271746 2.204312 2.230270 2.242083 2.264519 2.194403 2.171665 2.236958 2.201351 2.303154" and the actual values being "2.313525 2.226783 2.224624 2.319442 2.372111 2.326302 2.490723 2.447551 2.494857 2.562639"

#7
```{r}
viewers.model$ttable

```
Looking at the estimated coefficients of the autoregressive and moving average pieces, the only one that seems to be decreasing is the moving average coefficient of the model, and as such, the viewership may have moments with a decrease from one episode to another, but will over see an increase in viewership throughout each season.

#8
```{r}
pred.Season <- max(viewerData$ShowNum) + seq(1, 10, by=1)
Xpred <- predict(X, newx=pred.Season)
my.for <- sarima.for(viewerData$Viewers, p=2, d=0, q=0, P=0, D=1, Q=1, S=10, n.ahead=10, newxreg=Xpred)
my.for$pred
```
Looking at the predictions, there appears to still be a steady increase in viewership across the season showing that season 8 ends successfully and so a season 9 would be expected to have similar results.
